Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-05-15 Paper.Rmd
adding asym sens description : not completed yet

### Asymmetric sensitivity and copula structure in populations

<!--Another model, distinct from the above, Dan and Shya to discuss-->

@@

Figure \ref{fig_skewness_ff} shows that averaged skewness is much higher than expected if the aphid first flight data came from a normal copula. Here, $p$ value indicates the fraction of surrogate-skewness which are greater than the averaged skewness of the given data (red line in Figure \ref{SM-fig_skewness_ff}). As $p<0.05$, we reject $H_{0}$ in support for upper tail dependence.

@@

    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_K_effect}]

>>>>

### Asymmetric sensitivity and copula structure in populations
<!--Another model, distinct from the above, Dan and Shya to discuss-->
Here we intend to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) beyond a certain expectation threshold. To test this assymetric sensitivity, we constructed a two habitat patch model for population $P_{i} (i=1,2)$ as follows;

\begin{equation}\label{eq.AS}
P_{i}(t+1) = b P_{i}(t) + g(\epsilon_{i}(t)) + h(\delta_{i}(t))
\end{equation}

$b$ was the linear growth rate. Global environmental fluctuation function $g(\epsilon)$ can take up any one form between $g_{1}$ and $g_{2}$ (Eq. \ref{eq.AS_g}) where $\epsilon=(\epsilon_{1}, \epsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\epsilon}$.

\begin{align}
g(\epsilon) = \begin{cases} 
g_{1} = \begin{cases}
\epsilon & \text{if $\epsilon< 0$}; \\
0 & \text{otherwise}.
\end{cases}\\ \\
g_{2} = \begin{cases}
\epsilon & \text{if $\epsilon> 0$}; \\
0 & \text{otherwise}\label{eq.AS_g}.
\end{cases}
\end{cases} \text{where } \epsilon = N(\textbf{0},\Sigma_{\epsilon})\;\ \text{with} \;\ \Sigma_{\epsilon} = \begin{bmatrix} 1 & r \\ r & 1 \end{bmatrix}
\end{align}

Local environmental fluctuation function $h(\delta)$ varied according to Eq. \ref{eq.AS_h} where $a$ was the scaling parameter between $0$ to $1$. $\delta=(\delta_{1}, \delta_{2})$ was an iid across $i=(1,2)$ and time with mean zero and covariance matrix $\Sigma_{\delta}$.

\begin{equation}\label{eq.AS_h}
h(\delta) = a \delta = aN(\textbf{0},\Sigma_{\delta})\;\ \text{with} \;\ \Sigma_{\delta} = \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}
\end{equation}

We simulated our model (\ref{eq.AS}) for different values of $b$ with $a=0.2$, $r=0.8$. $N=10000$ points were generated from each local and global fluctuations to apply on the initial populations and then after simulation we kept only last $1000$ populations to plot the copula for a specific $b$ and a specific functional form of $g$. Fig. \ref{SM-fig_ms_asym_b_0.1} and Fig. \ref{SM-fig_ms_asym_b_0.9} show these results. We also carried out model selection and non-parametric statistical analysis with each copula of those figures and summarized our results in Tables \ref{SM-tab_summary_stat_asym_sens} - \ref{SM-tab_stat_npa_asym_8}. Table \ref{SM-tab_summary_stat_asym_sens} shows that according to our model selection approach for $b=\pm0.1$, non-normal copula (with higher AIC weights) were always best-fit with some asymetric tail dependence.

@@

Figure \ref{SM-fig_skewness_ff} shows that averaged skewness is much higher than expected if the aphid first flight data came from a normal copula. Here, $p$ value indicates the fraction of surrogate-skewness which are greater than the averaged skewness of the given data (red line in Figure \ref{SM-fig_skewness_ff}). As $p<0.05$, we reject $H_{0}$ in support for upper tail dependence.

@@

    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{SM-fig_risk_K_effect}]

### 2018-05-15 SupportingInformation.Rmd
adding asym sens description : not completed yet

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/SoilCNraw.pdf}
\includegraphics[width=3 in]{./Results/SoilCNcop.pdf}
\caption[\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data]{\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data. Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n = `r nrow(d)` $ is the number of sampling locations (see Sec. \ref{Model_selection}).\label{fig_raw_cop_soilCN}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/BMR_results/rawdataplot_BMR_vs_lnmass.pdf}
\includegraphics[width=3 in]{./Results/BMR_results/copula_BMR_vs_lnmass.pdf}
\caption[\textbf{(A)} Plot of the raw BMR vs. ln(mass) data \textbf{(B)} Copula plot for BMR dataset]{\textbf{(A)} Plot of the raw BMR vs. ln(mass) data \textbf{(B)} Copula plot for BMR dataset. Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_BMR)`$ is the number of contrasts (see Sec. \ref{Model_selection}).\label{fig_raw_cop_BMR}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}
\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
\caption[\textbf{(A)} Plot of the raw AVG.ltoe.IV vs. AVG.humerus phylogenetic traits. \textbf{(B)} Copula plot of AVG.ltoe.IV vs. AVG.humerus phylogenetic traits]{\textbf{(A)} Plot of the raw AVG.ltoe.IV vs. AVG.humerus phylogenetic traits. \textbf{(B)} Copula plot of AVG.ltoe.IV vs. AVG.humerus phylogenetic traits. Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_contrasts_67)`$ is the number of contrasts (see Sec. \ref{Model_selection}). \label{fig_raw_cop_AVG_humerus_ltoe}}
\end{center}
\end{figure}

@@

## Asymmetric sensitivity and copula structure in populations

<!--Another model, distinct from the above, Dan and Shya to discuss-->
```{r asym_sens_b_variation, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
set.seed(seed)
#-----------------------------------------------------------------------------------
source("OurBiCopSelect.R")
source("bivfunctionplot.R")
source("asym_sensitivity.R")
#-----------------------------------------------------------------------------------
# This function takes 
# Input :
#       mg : global noise matrix (dim = N by 2)
#       ml : local noise matrix (dim = N by 2)
#       a : a number : local noise coeeficient
#       b : growth factor of pop
#       decrease : logical (controls the nature of global noise)
#       numkeep_last : number of points you want to keep for the pop matrix from the end
#       ploton : logical for optional plot

@@

summary_stat<-as.data.frame(matrix(NA,7,8),stringsAsFactors = F) # Table for model selection summary
rownames(summary_stat)<-c("bestcop_AICw","N_AICw","relLTdep_AICw","relUTdep_AICw","p_CvM","p_KS","Numboot_success")
#bg <- intToUtf8(0x03B2) 
cnm<-c("b_0.1_g~1~","b_0.1_g~2~","b_-0.1_g~1~","b_-0.1_g~2~","b_0.9_g~1~","b_0.9_g~2~","b_-0.9_g~1~","b_-0.9_g~2~")
#cname<-paste0(bg,cnm)
colnames(summary_stat)<-cnm

@@

  summary_stat[1,ic]<-max(res$InfCritRes$AICw)
  summary_stat[2,ic]<-res$InfCritRes$AICw[1]
  summary_stat[3,ic]<-res$relLTdep_AICw
  summary_stat[4,ic]<-res$relUTdep_AICw
  summary_stat[5,ic]<-res$GofRes_CvM
  summary_stat[6,ic]<-res$GofRes_KS
  summary_stat[7,ic]<-res$Numboot_success

@@

![results for model selection \label{fig_ms_asym_b_0.1}](./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf){ width=40%, height=40% }

![results for model selection \label{fig_ms_asym_b_0.9}](./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf){ width=40%, height=40% }

@@

\newpage
```{r tab_stat_npa_asym,echo=F,results="asis"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")

for(i in c(1:8)){
  temp<-make_table_stat_fn(data=tab_asym[[i]])
  temp2<-knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[i]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence."))
  print(temp2)
  cat("\\newpage")
}
```

>>>>

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/SoilCNraw.pdf}
\includegraphics[width=3 in]{./Results/SoilCNcop.pdf}
\caption[\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data]{\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data. Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n = `r nrow(d)` $ is the number of sampling locations (see \nameref{MT-Model_selection} of manuscript).\label{fig_raw_cop_soilCN}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/BMR_results/rawdataplot_BMR_vs_lnmass.pdf}
\includegraphics[width=3 in]{./Results/BMR_results/copula_BMR_vs_lnmass.pdf}
\caption[\textbf{(A)} Plot of the raw BMR vs. ln(mass) data \textbf{(B)} Copula plot for BMR dataset]{\textbf{(A)} Plot of the raw BMR vs. ln(mass) data \textbf{(B)} Copula plot for BMR dataset. Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_BMR)`$ is the number of contrasts (see \nameref{MT-Model_selection} of manuscript).\label{fig_raw_cop_BMR}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}
\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
\caption[\textbf{(A)} Plot of the raw AVG.ltoe.IV vs. AVG.humerus phylogenetic traits. \textbf{(B)} Copula plot of AVG.ltoe.IV vs. AVG.humerus phylogenetic traits]{\textbf{(A)} Plot of the raw AVG.ltoe.IV vs. AVG.humerus phylogenetic traits. \textbf{(B)} Copula plot of AVG.ltoe.IV vs. AVG.humerus phylogenetic traits. Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_contrasts_67)`$ is the number of contrasts (see \nameref{MT-Model_selection} of manuscript). \label{fig_raw_cop_AVG_humerus_ltoe}}
\end{center}
\end{figure}

@@

<!--
## Asymmetric sensitivity and copula structure in populations
           ----moved to Paper.Rmd----
-->

```{r asym_sens_b_variation, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
set.seed(seed)
#-----------------------------------------------------------------------------------
source("OurBiCopSelect.R")
source("bivfunctionplot.R")
source("asym_sensitivity.R")
#-----------------------------------------------------------------------------------
# This function takes 
# Input :
#       mg : global noise matrix (dim = N by 2)
#       ml : local noise matrix (dim = N by 2)
#       a : a number : local noise coeeficient
#       b : growth factor of pop
#       decrease : logical (controls the nature of global noise)
#       numkeep_last : number of points you want to keep for the pop matrix from the end
#       ploton : logical for optional plot

@@

summary_stat<-as.data.frame(matrix(NA,7,8),stringsAsFactors = F) # Table for model selection summary
rownames(summary_stat)<-c("bestcop_AICw","N_AICw","relLTdep_AICw","relUTdep_AICw","p_CvM","p_KS","BS_success%")
#bg <- intToUtf8(0x03B2) 
cnm<-c("b_0.1_g~1~","b_0.1_g~2~","b_-0.1_g~1~","b_-0.1_g~2~","b_0.9_g~1~","b_0.9_g~2~","b_-0.9_g~1~","b_-0.9_g~2~")
#cname<-paste0(bg,cnm)
colnames(summary_stat)<-cnm

@@

  summary_stat[1,ic]<-max(res$InfCritRes$AICw)
  summary_stat[2,ic]<-res$InfCritRes$AICw[1]
  summary_stat[3,ic]<-res$relLTdep_AICw
  summary_stat[4,ic]<-res$relUTdep_AICw
  summary_stat[5,ic]<-res$GofRes_CvM
  summary_stat[6,ic]<-res$GofRes_KS
  summary_stat[7,ic]<-(res$Numboot_success/res$Numboot)*100

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
\caption[asym sensitivity results : copula plot]{asym sensitivity results : copula plot \label{fig_ms_asym_b_0.1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf}
\caption[asym sensitivity results : copula plot]{asym sensitivity results : copula plot \label{fig_ms_asym_b_0.9}}
\end{center}
\end{figure}

@@

```{r read_tab_stat_npa_asym,echo=F,results="asis"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")

#for(i in c(1:8)){
#  temp<-make_table_stat_fn(data=tab_asym[[i]])
#  temp2<-knitr::kable(temp, booktabs = T, 
#             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
#      caption =paste(cnm[i]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence."))
#  print(temp2)
#  cat("\\newpage")
#}
```

```{r tab_stat_npa_asym_1,echo=F,results="markup"}
temp<-make_table_stat_fn(data=tab_asym[[1]])
knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[1]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_1}"))
```

\newpage
```{r tab_stat_npa_asym_2,echo=F,results="markup"}
temp<-make_table_stat_fn(data=tab_asym[[2]])
knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[2]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_2}"))
```

\newpage
```{r tab_stat_npa_asym_3,echo=F,results="markup"}
temp<-make_table_stat_fn(data=tab_asym[[3]])
knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[3]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_3}"))
```

```{r tab_stat_npa_asym_4,echo=F,results="markup"}
temp<-make_table_stat_fn(data=tab_asym[[4]])
knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[4]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_4}"))
```

```{r tab_stat_npa_asym_5,echo=F,results="markup"}
temp<-make_table_stat_fn(data=tab_asym[[5]])
knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[5]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_5}"))
```

```{r tab_stat_npa_asym_6,echo=F,results="markup"}
temp<-make_table_stat_fn(data=tab_asym[[6]])
knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[6]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_6}"))
```

```{r tab_stat_npa_asym_7,echo=F,results="markup"}
temp<-make_table_stat_fn(data=tab_asym[[7]])
knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[7]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_7}"))
```

```{r tab_stat_npa_asym_8,echo=F,results="markup"}
temp<-make_table_stat_fn(data=tab_asym[[8]])
knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[8]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_8}"))
```
